Bare Demo of IEEEtran.cls\\ for IEEE Conferences

### 2016-11-30 paper.tex
Comments to conclusions and errors in Abstract

Learning to program is often regarded as a difficult task, courses are 
problem-solving intensive and often students feel frustrated when doing programming
exercises. Instructors aim to design learning activities that result
challenging to students, but not to the extreme as to frustrate them. When
selecting an appropiate programing excercise, experienced instructors gauge a
student’s affective state and skills to then assign an activity with the  appropriate level
of difficulty. This work is focused on the prediction of the affective states
of programmers with different levels of expertise when learning a new
programming language. For this, an interactive web-based programming platform
is proposed. The platform is designed to collect data from the students'
interaction for data analysis. Current work is focused on the prediction of
affective states using non-obtrusive sensors. Specifically, the proposed method uses data
gathered from both keyboard and mouse dynamics. The data set was generated
as students learned the basics of Python programming. Feature vectors 
were generated with 39 features representing the dynamics of the interaction. 
In total 149 feature vectors were generated and used as input of four different
classification algorithms: Naïve Bayes, J-48, k-NN, and Artificial Neural
Networks. These classification models are then compared by their performances
when classifying learners into five different affective states: boredom,
frustration, distraction, relaxation and engagement. Each of the classification
models achieved an accuracy of around 75% with a kappa of around 0.45, and, in
average, J-48 was the better algorithm with an accuracy of 75.97%. Results show
that the data gathered from the non-obtrusive sensors can successfully be used
as another input to classification models in order to predict an individual's
affective states.

@@

The participants' interaction generated a total of 142 feature vectors one for each
successfully completed exercise. The affective states reported by learners after
completing each exercise was grouped in three classes: ‘Yes’, ‘Neutral’ and
‘No’. The class distribution of the emotions reported is shown in Figure 6. This
results show that the most common emotions were flow/engagement (72\%) and
relaxation (61\%) while few learners reported distraction (5\%), frustration
(8\%) and boredom (2\%). This distribution is different from what was reported
by D’Mello et al. \cite{bixler2013detecting} and Rodrigo et al. \cite{rodrigo2009affective}.
Although Flow/Engagement was
also the predominant class, the distribution is skewed to the first two. There
are some possible reasons for this; first the majority of students had
experience in programming, so learning new one is not that problematic, a second
reason could be the freedom users had to abandon the activities, perhaps
frustrated or bored learners simply quit the tutorial. A high number of learners
(49\%) did not completed any exercise and the once who completed more where also
the more experienced.

Classifiers where implemented using the software Rapid Miner using a feature
selection module as part of the process. A genetic algorithm was used in the
user selection module. In the end, the subset for the frustration classifier
consists of 11 features, and the subset for the boredom classifier consists of
13 features. The performance of the classifiers is shown in Table 1.

The results obtained with this method are satisfactory, considering a free-text
was used. The accuracies and kappa coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk (2007). In the case of the kappa coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of kappa were close to 0.5, is important to consider the values of kappa
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\% marginally better than the others, and an accepted kappa
statistic for the kind of problem. Artificial neural network give the highest
accuracy in the classification of frustration.

\section{Conclusion}
The aim of this research was to evaluate the use of keyboard and mouse dynamics 
as an appropriate sensory input for an affective recognition system. The context
of use would be learning environment for programmers. In particular for learning
activities consisting in writing short programs interactively. An experimental
approach was adopted with this aim. The proposed method in this work obtained
satisfactory results. It is usual for a classification method based on free-text
to obtain an accuracy and kappa below their counterparts of fixed-text, and in this
work the results obtained were similar to those works applied to fixed-text 
dynamics. A possible explanation of this would be the addition of the mouse 
dynamics features and the additional preprocessing performed on the feature vectors. 
While these are promising results further experiments are needed. Other experiments
should focus on novice programmers.

>>>>

Learning to program is often regarded as a difficult task, courses are 
problem-solving intensive and often students feel frustrated when doing programming
exercises. Instructors aim to design learning activities that result
challenging to students, but not to the extreme as to frustrate them. When
selecting an appropiate programing excercise, experienced instructors gauge a
student’s affective state and skills to then assign an activity with the  appropriate level
of difficulty. This work is focused on the prediction of the affective states
of programmers with different levels of expertise when learning a new
programming language. For this, an interactive web-based programming platform
is proposed. The platform is designed to collect data from the students'
interaction for data analysis. Current work is focused on the prediction of
affective states using non-obtrusive sensors. Specifically, the proposed method uses data
gathered from both keyboard and mouse dynamics. The data set was generated
as students learned the basics of Python programming. Feature vectors 
were generated with 39 features representing the dynamics of the interaction. 
In total 149 feature vectors were generated and used as input of four different
classification algorithms: Naïve Bayes, J-48, k-NN, and Artificial Neural
Networks. These classification models are then compared by their performances
when classifying learners into five different affective states: boredom,
frustration, distraction, relaxation and engagement. Each of the classification
models achieved an accuracy of around 75\% with a kappa of around 0.45, and, in
average, J-48 was the better algorithm with an accuracy of 75.97\%. Results show
that the data gathered from the non-obtrusive sensors can successfully be used
as another input to classification models in order to predict an individual's
affective states.

@@

The participants' interaction generated a total of 142 feature vectors one for each
successfully completed exercise. The affective states reported by learners after
completing each exercise was grouped in three classes: ‘Yes’, ‘Neutral’ and
‘No’. The class distribution of the emotions reported is shown in Figure 6. This
results show that the most common emotions were flow/engagement (72\%) and
relaxation (61\%) while few learners reported distraction (5\%), frustration
(8\%) and boredom (2\%). This distribution is different from what was reported
by D’Mello et al. \cite{bixler2013detecting} and Rodrigo et al. \cite{rodrigo2009affective}.
Although Flow/Engagement was
also the predominant class, the distribution is skewed to the first two. There
are some possible reasons for this; first the majority of students had
prior experience in programming, so learning new one is not that problematic, a second
reason could be the freedom users had to abandon the activities, perhaps
frustrated or bored learners simply quit the tutorial. A high number of learners
(49\%) did not completed any exercise and the once who completed more where also
the more experienced.

Classifiers where implemented using the software Rapid Miner using a feature
selection module % which one? - JJ
as part of the process. A genetic algorithm was used in the
user selection module. % Which one? You really have to be _way_ more
                       % specific in this conference - JJ
In the end, the subset for the frustration classifier
consists of 11 features, and the subset for the boredom classifier consists of
13 features. The performance of the classifiers is shown in Table 1.

%Start with a reference to the Table (and check numbering) - JJ
The results obtained with this method are satisfactory, considering a free-text
was used. % don't get this. What free text? A program? - JJ
The accuracies and kappa coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk (2007). %A reference!!! - JJ
 In the case of the kappa coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of kappa were close to 0.5, is important to consider the values of kappa
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\%, which is marginally better than the others, and an accepted kappa
statistic for the kind of problem. The artificial neural network % Which one? - JJ
give the highest
accuracy in the classification of frustration.

\section{Conclusion}
The aim of this research was to evaluate the use of keyboard and mouse dynamics 
as an appropriate sensory input for an affective recognition system. The context
of use would be learning environment for programmers. In particular for learning
activities consisting in writing short programs interactively. An experimental
approach was adopted with this aim. The proposed method in this work obtained
satisfactory results. It is usual for a classification method based on free-text
to obtain an accuracy and kappa below their counterparts of fixed-text, and in this
work the results obtained were similar to those works applied to fixed-text 
dynamics. A possible explanation of this would be the addition of the mouse 
dynamics features and the additional preprocessing performed on the feature vectors. 
While these are promising results further experiments are needed. Other experiments
should focus on novice programmers.
% But you have to comment on how this addresses the problem. Which
% should be in the title that is not there yet - JJ
% You also have to draw a _hard_ conclusion. Is programming using free
% text better than what? Typing a fixed text?
